# Data sources
-Dataset1 ("compas-scores-two-years.csv"):Broward County Clerk’s Office, Broward County Sherrif's Office, Florida Department of Corrections, ProPublica.
This database contains criminal history, jail and prison time, demographics and COMPAS risk scores for defendants from Broward County.

-Dataset2 ("12 Variables NYPD Arrest Data.xlsx"):Data collected by the New York City Police Dept. We selected 12 Variables that are relevant to our project:
$1.ARREST KEY:$ Unique identifier for the record.  

$2.ARREST DATE:$ Date when the incident occurred.  

$3.ARREST BORO:$ Borough of arrest within four categories:
B(Bronx), S(Staten Island), K(Brooklyn), M(Manhattan), Q(Queens).  

$4.AGE GROUP:$ Perpetrator age within a certain category.  

$5.PERP RACE:$ Perpetrator race description.  

$6.PERP SEX:$Perpetrator sex description.  

$7.OFNS DESC:$ Type of crime committed.  

$8.ARREST PRECINCT:$geographical area patrolled by a police force.  

$9.X COORD CD:$ The x coordinate of the location where the incident occurred.  

$10.Y COORD CD:$ The y coordinate of the location where the incident occurred.  

$11.Latitude:$ The latitude of the location where the incident occurred.  

$12.Longitude:$ The longitude of the location where the incident occurred.  


-Dataset3 ("geo_export_9f1de9c6-45df-4c9a-8939-1843acf3378c.shp"):The metadata for New York City Political, Administrative and Census geographies. The shape file was generated from the 21A release of the Department of City Planning's LION file.We downloaded this from NYC government website.
```{r}
library(dplyr)
library(ggplot2)
library(lattice)
library(caret)
library(ggplot2)
library(forcats)
library(treemap)
library(vcd)
library(grid)
library(ggmosaic)
library(GGally)
library(RColorBrewer)
library(lubridate)
library(hrbrthemes)
library(viridis)
library(sp)
library(sf)
library(rgdal)
library(readxl)
library(plyr)
```

```{r}
#Dataset 1
raw_data <- read.csv("/Users/jessie/Desktop/compas-scores-two-years.csv")
head(raw_data)

#Dataset 2
dt<-readxl::read_excel("/Users/jessie/Desktop/STAT 5702/Final Project/12 Variables NYPD Arrest Data.xlsx")
head(dt)

#Dataset 3
shpdt<-st_read("/Users/jessie/Downloads/Police Precincts/geo_export_9f1de9c6-45df-4c9a-8939-1843acf3378c.shp")
```